---
title: "NHSD Application"
author: "GeoC"
date: "15/04/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r readIn, include = FALSE}
 
library(dplyr)
library(readr)

report_date <- "2019-04-12"

#list claning Q1_18 two installments
MR1418_01_ListCleaningCurrentStatusReport_50000_20180918_RMD <- read_csv("~/NHSD/Q1_2018_Data_Return/MR1418_01_ListCleaningCurrentStatusReport_50000_20180918_RMD.csv", 
    col_names = FALSE, col_types = cols(X1 = col_character(), X2 = col_character()))

MR1418_02_ListCleaningCurrentStatusReport_37263_20180918_RMD <- read_csv("~/Documents/NHSD_HES/PastApplications/MR1418_ListCleaningReport_20180920/MR1418_02_ListCleaningCurrentStatusReport_37263_20180918_RMD.csv", 
    col_names = FALSE, col_types = cols(X1 = col_character(), X2 = col_character()))

lc <- rbind.data.frame(MR1418_01_ListCleaningCurrentStatusReport_50000_20180918_RMD, MR1418_02_ListCleaningCurrentStatusReport_37263_20180918_RMD)

#join and make readable
lc <- lc %>% select(X1, X2)
names(lc) <- c("participant_id", "nhs_no")

#latest flagging Q2_18
MR1418_01_FlaggingCurrentStatusReport_12886_20181217_RMD <- read.csv("~/NHSD/Q2_2018_Data_Return/RAW/MR1418_01_FlaggingCurrentStatusReport_12886_20181217_RMD.csv")
#fl <- MR1418_01_FlaggingCurrentStatusReport_12886_20181217_RMD %>% select(Supplied.Member.Number, Latest.NHS.Number..New.) %>% distinct()
#names(fl) <- c("participant_id", "nhs_no")

#remove the twin participant_ids of DUPLICATE and then distinctify the amended/cancelled multiples 
MR1418_01_FlaggingCurrentStatusReport_12886_20181217_RMD %>% filter(!grepl("DUPLICATE", Cancelled.Event...Amended.Event..Duplicated.Patient.Notification )) %>% select(Supplied.Member.Number, Latest.NHS.Number..New.) %>% distinct() -> fl2
names(fl2) <- c("participant_id", "nhs_no")
fl2$participant_id <- as.character(fl2$participant_id)
lc$participant_id <- as.character(lc$participant_id)
participants_old <- rbind.data.frame(fl2, lc, stringsAsFactors = FALSE)

#output of elibible flag sql script. 
AllGelEligibility_12_04_19 <- read_csv("~/SQL/CSV_Outputs/AllGelEligibility_12_04_19.csv",
col_types = cols(participant_id = col_character()))
# View(AllGelEligibility_12_04_19)

#output of all registered participants with their address and ods codes.
AllGelRegODS_12_04_19 <- read_csv("~/SQL/CSV_Outputs/AllGelRegODS_12_04_19.csv",
col_types = cols(participant_id = col_character()),
na = "empty")
# View(AllGelRegODS_12_04_19)

#output of all withdrawals script
AllGelWRD_12_04_19 <- read_csv("~/SQL/CSV_Outputs/AllGelWRD_12_04_19.csv",
col_types = cols(participant_identifiers_id = col_character()),
na = "empty")

#old application
MR1418_01_108786 <- read_csv("~/Downloads/MR1418_01_10786.csv", col_names=FALSE, col_types= cols(X2  = col_character()), na="empty" )

#old_applicationCLLs
CLLAllHist <- participants_old %>% 
  filter(grepl("-",participant_id )) 


#old_application_mainProgramme
MainProgrammeAllHist <- participants_old %>% 
    filter(!participant_id %in% CLLAllHist$participant_id) %>% 
    filter(nchar(participant_id) > 8)

#old_application_RDPilot
RDPilotAllHist <- participants_old %>% filter(nchar(participant_id) < 6)  

#View(AllGelWRD_12_04_19)
#load("/Users/georgiachan/NHSD/Q2_2018_Data_Return/RData/valid_participants_Q2_2018.RData")
#dim(validpartids)


```

## Overview

Numbers valid on `r report_date`

**Registered**    
  All registered DAMS participants = **`r dim(AllGelRegODS_12_04_19)[1]`** (includes test participants)  
  All registered DDF participants = **`r dim(AllGelEligibility_12_04_19)[1]`** 
  
**Withdrawn**  
  Total of fully withrdrawn participants = **`r dim(AllGelWRD_12_04_19 %>% filter((grepl("full", withdrawal_option_id))|(grepl("FULL", withdrawal_option_id))))[1]`**

**List cleaning plus latest flagging**  
  The entire list of participants that have been traced by NHSD to this date = **`r dim(participants_old)[1]`**  
  Includes RD Pilot, First Phase of CLL and DAMS Main Programme. 
 
  + **Main programme**  
      Total = **`r dim(MainProgrammeAllHist)[1]`**
  
  + **CLL participants (1st application) **  
      CLL cohort lives outside DAMS (apart from validation cohort that has no other PID attached to be traceable). Secondary data is requested since Q1_18.
      All CLL participants traced in the existing NHSD cohort:  
      Total = **`r  dim(CLLAllHist)[1]`**  (out of  `r dim(MR1418_01_108786 %>% filter(grepl("-", X2)))[1]`). 
  
  + **Rare Disease pilot**
      Rare disease pilot lives outside DAMS (although there are participants recruited in the main programme as well that were flagged as duplicates from NHSD since they have the same NHS number). Their total number and consent status is frozen. Total = **`r dim(RDPilotAllHist)[1]`**
    
    
## Eligibility exclusions (drawn from DDF). 
Exemptions from application to secondary data:

```{r eligibility, include = FALSE}
cohort_ods <- c("TRX", "BCN", "MAJ", "PHA", "ALL", "NEQ")
#Participants in cohorts
GelCohortPids <- AllGelRegODS_12_04_19 %>% filter(ods %in% cohort_ods) %>% select(participant_id)

#eligible from sample presence
eligible.pid <- AllGelEligibility_12_04_19 %>% filter(eligible==TRUE)
non.eligible.pid <- AllGelEligibility_12_04_19 %>% filter(eligible==FALSE)

#non eligible from wrong type of consent (age) 
load("./consent.RData")
child.consent <- d %>% filter(require_reconsent==TRUE) 
cd <- child.consent %>% select(participant_identifiers_id, require_reconsent) 
names(cd) <- c("participant_id", "require_reconsent")

#eligible_from_valid <- validpartids %>% filter(participant_id %in% eligible.pid$participant_id)
#(dim(validpartids) - dim(eligible_from_valid))/dim(validpartids)=
#(dim(validpartids) - dim(eligible_from_valid))/dim(validpartids)
sum(GelCohortPids$participant_id %in% MainProgrammeAllHist$participant_id)

```

* **Cohorts**

    ODS codes of cohorts with consent prohibiting linkage to secondary datasets:  
      **TRX**, **BCN**, **MAJ**, **PHA**, **ALL**, **NEQ**.
    Total = `r dim(GelCohortPids)[1]`
    
* **Consent**
  + All fully withdrawn participants. Total = **`r dim(AllGelWRD_12_04_19 %>% filter((grepl("full", withdrawal_option_id))|(grepl("FULL", withdrawal_option_id))))[1]`**
  + Participants who are 16+ but on chid concent form + participants whose form will be invalid before next application. **`r dim(child.consent)[1]`**

* **Sample Non-Eligibility Criteria**

  + **Cancer** - All cancer cases without a samples not sent notification.
  + **Rare D** - All rare disease cases without DNA sample.  

    Total eligible (DDF) = **`r dim(eligible.pid)[1]`**  
    Total non eligible = **`r dim(non.eligible.pid)[1]`**
    
```{r final, include = FALSE }
participants_newInc <- eligible.pid %>% 
  filter(!participant_id %in% participants_old$participant_id)


#new list should exclude 
participants_newExcExc <- AllGelEligibility_12_04_19 %>% 
  filter(!participant_id %in% participants_old$participant_id) %>% #exclude old participants
  filter(!participant_id %in% GelCohortPids$participant_id) %>% #exclude cohorts
  filter(!participant_id %in% cd$participant_id) %>%  #filter consents
  filter(participant_id %in% eligible.pid$participant_id) #filter sample exclusions
  
participants_newIncExc <- AllGelEligibility_12_04_19 %>% 
  filter(!participant_id %in% participants_old$participant_id) %>% #exclude old participants
  filter(!participant_id %in% GelCohortPids$participant_id) %>% #exclude cohorts
  filter(!participant_id %in% cd$participant_id) #filter consents

ApplicationNewExcExc <- AllGelRegODS_12_04_19 %>%
  filter(participant_id %in% participants_newExcExc$participant_id)
ApplicationNewIncExc <- AllGelRegODS_12_04_19 %>%
  filter(participant_id %in% participants_newIncExc$participant_id)

participants_wdrIncExc <- c(non.eligible.pid$participant_id, GelCohortPids$participant_id, cd$participant_id)
participants_wdrExcExc <- c(GelCohortPids$participant_id, cd$participant_id)
  
save(ApplicationNewExcExc, ApplicationNewIncExc, participants_wdrIncExc,participants_wdrExcExc, file= "./Application_Q3_18.RData" )
#eligible.pid %>% filter(!participant_id %in% MainProgrammeAllHist$participant_id ) -> newPids
```

## New Application

* **Main programme**  
    
    Total new without exclusions = **`r dim(participants_newIncExc)[1]`**  
    Total new with exclusions = **`r dim(participants_newExcExc)[1]`**  
    
* **CLL new**  
    Total = 16

* **Withdrawals**  

    Total withdrawasl with exclusions = **`r length(participants_wdrIncExc)`**  
    Total withdrawals without exclusions = **`r length(participants_wdrExcExc)`**    
 